---
title: "Conservation Genomics Report"
date: "`r params$report_date`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    self_contained: true
params:
  pop_stats:   NULL
  fst_mat:     NULL
  pop_q:       NULL
  ne_stats:    NULL
  gea_data:    NULL
  offset_data: NULL
  is_demo:     FALSE
  report_date: ""
  app_version: ""
---

```{css, echo=FALSE}
body { font-family: 'Segoe UI', Arial, sans-serif; font-size: 14px; color: #333; }
h1, h2, h3, h4 { color: #1B3A5C; }
h2 { border-bottom: 2px solid #2E75B6; padding-bottom: 6px; margin-top: 2em; }
h3 { color: #2E75B6; margin-top: 1.5em; }
.risk-badge {
  display: inline-block; padding: 3px 10px; border-radius: 12px;
  color: white; font-size: 0.85em; font-weight: 600;
}
.summary-grid { display: flex; gap: 16px; flex-wrap: wrap; margin: 16px 0; }
.summary-card {
  flex: 1 1 160px; padding: 14px 18px; border-radius: 8px;
  border-left: 4px solid #2E75B6; background: #f8f9fa;
  box-shadow: 0 1px 4px rgba(0,0,0,0.07);
}
.summary-card .value { font-size: 1.6em; font-weight: 700; color: #1B3A5C; }
.summary-card .label { font-size: 0.78em; color: #777; margin-top: 2px; }
table { width: 100%; border-collapse: collapse; font-size: 0.9em; }
thead th { background: #1B3A5C; color: white; padding: 8px 10px; text-align: left; }
tbody td { padding: 7px 10px; border-bottom: 1px solid #eee; }
tbody tr:nth-child(even) { background: #fafafa; }
.alert-demo { background: #FFF3CD; border: 1px solid #ffc107; border-radius:6px;
               padding: 10px 14px; margin-bottom: 16px; font-size: 0.88em; }
.section-note { font-size: 0.82em; color: #777; font-style: italic; margin-top: 6px; }
.unavailable { color: #aaa; font-style: italic; font-size: 0.9em; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(knitr)

# Bring app helper functions into scope
# (the report is rendered in a fresh env; source helpers explicitly)
fmt_r <- function(x, digits = 3) {
  ifelse(is.na(x), "—", formatC(round(x, digits), format = "f", digits = digits))
}

he_risk <- function(he) {
  if (is.na(he))   return("Unknown")
  if (he < 0.10)   return("Critical — very low diversity")
  if (he < 0.15)   return("High risk — low diversity")
  if (he < 0.20)   return("Moderate risk")
  if (he < 0.30)   return("Low risk — moderate diversity")
  return("Healthy — high diversity")
}

he_colour <- function(he) {
  if (is.na(he)) return("#AAAAAA")
  if (he < 0.10) return("#D32F2F")
  if (he < 0.15) return("#F57C00")
  if (he < 0.20) return("#FBC02D")
  if (he < 0.30) return("#388E3C")
  return("#1565C0")
}

ne_risk <- function(ne) {
  if (is.null(ne) || is.na(ne)) return("Not estimated")
  if (ne < 50)  return("Critical")
  if (ne < 100) return("High risk")
  if (ne < 500) return("Concern")
  return("Viable")
}

df        <- params$pop_stats
fst_mat   <- params$fst_mat
pop_q     <- params$pop_q
ne_stats  <- params$ne_stats
gea_data  <- params$gea_data
offset_df <- params$offset_data
is_demo   <- isTRUE(params$is_demo)
```

---

> **Generated by PopGen Map v`r params$app_version`** &nbsp;|&nbsp; `r params$report_date`
> Open-source conservation genomics platform — [github.com/csunbird/population-genomics-map](https://github.com/csunbird/population-genomics-map)

```{r demo-banner, results='asis'}
if (is_demo) {
  cat('<div class="alert-demo">⚠️ <strong>Demo dataset:</strong> This report was generated from simulated data. Values are illustrative only — not real genotype data for any species.</div>\n')
}
```

---

# 1. Executive Summary

```{r summary-cards, results='asis'}
n_pops  <- nrow(df)
n_inds  <- sum(df$n, na.rm = TRUE)
mean_he <- mean(df$He, na.rm = TRUE)
at_risk <- sum(df$He < 0.20, na.rm = TRUE)

card <- function(value, label, col = "#2E75B6") {
  sprintf(
    '<div class="summary-card" style="border-left-color:%s;">
       <div class="value" style="color:%s;">%s</div>
       <div class="label">%s</div>
     </div>',
    col, col, value, label
  )
}

cat('<div class="summary-grid">\n')
cat(card(n_pops, "Populations", "#1B3A5C"))
cat(card(n_inds, "Individuals", "#2E75B6"))
cat(card(fmt_r(mean_he), "Mean He", "#388E3C"))
cat(card(at_risk, "At-risk populations (He < 0.20)",
         if (at_risk > 0) "#D32F2F" else "#388E3C"))
cat('</div>\n')
```

```{r risk-overview, results='asis'}
risk_counts <- table(sapply(df$He, he_risk))
cat("**Genetic health overview:**\n\n")
for (risk_label in names(risk_counts)) {
  col <- he_colour(switch(risk_label,
    "Critical — very low diversity"    = 0.05,
    "High risk — low diversity"        = 0.12,
    "Moderate risk"                         = 0.17,
    "Low risk — moderate diversity"    = 0.25,
    "Healthy — high diversity"         = 0.35,
    NA_real_
  ))
  cat(sprintf(
    '- <span class="risk-badge" style="background:%s;">%s</span> &nbsp; **%d** population%s\n',
    col, risk_label, risk_counts[[risk_label]],
    if (risk_counts[[risk_label]] == 1) "" else "s"
  ))
}
```

---

# 2. Per-Population Diversity Statistics

```{r stats-table}
display <- data.frame(
  Population = df$population,
  n          = df$n,
  SNPs       = if ("n_loci" %in% names(df)) df$n_loci else NA_integer_,
  He         = fmt_r(df$He),
  Ho         = fmt_r(df$Ho),
  `π`   = fmt_r(df$pi),
  F          = fmt_r(df$inbreeding_f, 3),
  Risk       = sapply(df$He, he_risk),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

knitr::kable(display, align = c("l", "c", "c", "c", "c", "c", "c", "l"),
             caption = "Per-population genetic diversity metrics")
```

<p class="section-note">He = expected heterozygosity | Ho = observed heterozygosity | π = nucleotide diversity (Nei 1978, diploid-corrected) | F = inbreeding coefficient (He − Ho) / He. Conservation thresholds (He) calibrated on microsatellite data; SNP-based He typically runs lower — treat risk categories as indicative.</p>

---

# 3. Population Structure

## 3.1 Pairwise F~ST~ Matrix

```{r fst-section, results='asis'}
if (is.null(fst_mat) || nrow(fst_mat) < 2) {
  cat('<p class="unavailable">FST matrix not available for this dataset (requires ≥ 2 populations with sufficient SNPs).</p>\n')
} else {
  fst_display <- round(fst_mat, 3)
  fst_display[upper.tri(fst_display)] <- NA
  print(knitr::kable(fst_display, caption = "Pairwise Hudson F_ST (lower triangle; ratio-of-averages estimator)"))

  # Wright thresholds note
  cat('\n<p class="section-note">',
      'Wright (1978) guideline: F_ST &lt; 0.05 = little differentiation; ',
      '0.05–0.15 = moderate; 0.15–0.25 = great; &gt; 0.25 = very great differentiation.</p>\n')

  # Flag highly differentiated pairs
  pairs <- which(fst_mat > 0.15 & upper.tri(fst_mat), arr.ind = TRUE)
  if (nrow(pairs) > 0) {
    cat("\n**⚠️ Highly differentiated pairs (F_ST > 0.15):**\n\n")
    for (k in seq_len(nrow(pairs))) {
      cat(sprintf("- %s ↔ %s: F_ST = %.3f\n",
                  rownames(fst_mat)[pairs[k,1]],
                  colnames(fst_mat)[pairs[k,2]],
                  fst_mat[pairs[k,1], pairs[k,2]]))
    }
    cat("\nHigh F_ST between populations may indicate barriers to gene flow — ",
        "consider connectivity assessments or managed translocation programmes.\n")
  }
}
```

## 3.2 ADMIXTURE Ancestry

```{r admixture-section, results='asis'}
if (is.null(pop_q)) {
  cat('<p class="unavailable">ADMIXTURE Q-matrix not uploaded. To include ancestry proportions, load a Q-matrix CSV via the ADMIXTURE tab.</p>\n')
} else {
  k_cols <- grep("^K[0-9]+$", names(pop_q), value = TRUE)
  K      <- length(k_cols)
  cat(sprintf("Q-matrix uploaded: **K = %d** genetic clusters across **%d** populations.\n\n", K, nrow(pop_q)))

  admix_display <- pop_q[, c("population", k_cols), drop = FALSE]
  admix_display[, k_cols] <- round(admix_display[, k_cols], 3)
  print(knitr::kable(admix_display,
    caption = sprintf("Per-population mean ADMIXTURE ancestry proportions (K = %d)", K)))
}
```

---

# 4. Inbreeding

```{r inbreeding-section, results='asis'}
# F-ROH data is not passed to the export module (it runs as an independent tab).
# Show the HWE-based F coefficient from pop_stats, which is always available.
# Users wanting per-individual ROH data can view or export from the F-ROH tab.

inb_df <- df[, c("population", "inbreeding_f"), drop = FALSE]
inb_df$inbreeding_f <- fmt_r(inb_df$inbreeding_f, 3)
inb_df$flag <- ifelse(df$inbreeding_f > 0.1, "⚠️ Elevated", "")
names(inb_df) <- c("Population", "F", "Flag")
print(knitr::kable(inb_df, caption = "Inbreeding coefficient F = (He − Ho) / He"))

elevated <- sum(df$inbreeding_f > 0.1, na.rm = TRUE)
if (elevated > 0) {
  cat(sprintf("\n**%d population%s show elevated inbreeding (F > 0.10).** ",
              elevated, if (elevated == 1) "" else "s"))
  cat("Consider genetic rescue (managed introduction of individuals from genetically ",
      "distinct populations) or a supervised breeding programme.\n")
}

cat('\n<p class="section-note">',
    'F = (He − Ho) / He.  Positive values indicate excess homozygosity relative to Hardy-Weinberg expectations. ',
    'For recent-inbreeding analysis using runs of homozygosity (F-ROH), see the F-ROH tab in the app — ',
    'F-ROH captures inbreeding over the last ~10 generations and is a more powerful signal than F from HWE departure.</p>\n')
```

---

# 5. Effective Population Size (Ne)

```{r ne-section, results='asis'}
if (is.null(ne_stats) || nrow(ne_stats) == 0) {
  cat('<p class="unavailable">Ne estimation results not available. Run the Ne module in the app to include effective population size estimates.</p>\n')
} else {
  ne_display <- ne_stats
  if ("ne" %in% names(ne_display)) {
    ne_display$risk <- sapply(ne_display$ne, ne_risk)
    ne_display$ne   <- round(ne_display$ne)

    show_cols <- intersect(c("population", "ne", "ne_lower", "ne_upper", "risk"),
                           names(ne_display))
    ne_tbl <- ne_display[, show_cols, drop = FALSE]
    names(ne_tbl)[names(ne_tbl) == "ne"]       <- "Ne"
    names(ne_tbl)[names(ne_tbl) == "ne_lower"] <- "Ne (lower 95%)"
    names(ne_tbl)[names(ne_tbl) == "ne_upper"] <- "Ne (upper 95%)"
    names(ne_tbl)[names(ne_tbl) == "risk"]     <- "Risk"

    print(knitr::kable(ne_tbl,
      caption = "Effective population size (Ne) estimated by linkage disequilibrium method (Waples & Do 2008)"))

    critical <- sum(ne_display$ne < 50, na.rm = TRUE)
    concern  <- sum(ne_display$ne < 500 & ne_display$ne >= 50, na.rm = TRUE)

    if (critical > 0)
      cat(sprintf("\n**%d population%s at critical Ne < 50** (immediate extinction risk from genetic drift).\n",
                  critical, if (critical == 1) "" else "s"))
    if (concern > 0)
      cat(sprintf("**%d population%s at Ne < 500** (long-term viability concern).\n",
                  concern, if (concern == 1) "" else "s"))
    if (critical == 0 && concern == 0)
      cat("\nAll populations have Ne ≥ 500 — no immediate Ne-based viability concern.\n")
  }

  cat('\n<p class="section-note">IUCN/SSC thresholds: Ne &lt; 50 = immediate extinction risk; ',
      'Ne &lt; 100 = high risk; Ne &lt; 500 = long-term viability concern (Frankham et al. 2014).</p>\n')
}
```

---

# 6. Adaptive Potential

## 6.1 Genotype-Environment Association (GEA)

```{r gea-section, results='asis'}
# gea_data is a list: $rda_result (list from calc_rda) + $lfmm_result (data frame)
if (is.null(gea_data) ||
    (is.null(gea_data$rda_result) && is.null(gea_data$lfmm_result))) {
  cat('<p class="unavailable">GEA results not available. Upload environmental data and run the GEA module to identify candidate adaptive loci.</p>\n')
} else {
  # ── Partial RDA ─────────────────────────────────────────────────────────────
  if (!is.null(gea_data$rda_result)) {
    cat("### Partial RDA\n\n")
    rda_r <- gea_data$rda_result
    if (!is.null(rda_r$pct_constrained)) {
      cat(sprintf(
        "Environmental variables explain **%.1f%%** of total genetic variance ",
        rda_r$pct_constrained
      ))
      cat("after partialling out geographic effects (lat/lon as covariates).\n\n")
    }
    if (!is.null(rda_r$top_loci_idx) && length(rda_r$top_loci_idx) > 0) {
      top_n <- min(10L, length(rda_r$top_loci_idx))
      cat(sprintf("**Top %d candidate loci** (highest |RDA1 loading|):", top_n), "\n\n")
      cat(paste0("- Locus index ", head(rda_r$top_loci_idx, top_n), collapse = "\n"), "\n\n")
      cat('<p class="section-note">Locus indices refer to positions in the filtered biallelic-SNP allele-frequency matrix. ',
          'To obtain genomic coordinates or SNP IDs, cross-reference these indices with your source VCF file. ',
          'Filter the VCF to biallelic SNPs only (same QC as used during upload) and select rows by position.</p>\n\n')
    }
    if (!is.null(rda_r$env_vars)) {
      cat("**Environmental variables tested:**",
          paste(rda_r$env_vars, collapse = ", "), "\n\n")
    }
  }

  # ── LFMM2 ───────────────────────────────────────────────────────────────────
  if (!is.null(gea_data$lfmm_result)) {
    cat("### LFMM2\n\n")
    lfmm_r  <- gea_data$lfmm_result
    n_sig   <- sum(lfmm_r$significant, na.rm = TRUE)
    env_tab <- sort(table(lfmm_r$env_var[lfmm_r$significant]), decreasing = TRUE)
    cat(sprintf(
      "**%d loci** reached significance at FDR < 5%% (Benjamini-Hochberg correction).\n\n",
      n_sig
    ))
    if (n_sig > 0 && length(env_tab) > 0) {
      cat("**Significant loci per environmental variable:**\n\n")
      for (vn in names(env_tab)) {
        cat(sprintf("- %s: %d significant loci\n", vn, env_tab[[vn]]))
      }
      cat("\n")
    }
    gif_vals <- unique(round(lfmm_r$gif, 2))
    cat(sprintf("Genomic inflation factor (GIF): %s.  ",
                paste(gif_vals, collapse = ", ")))
    cat("GIF near 1.0 indicates well-calibrated p-values; GIF > 2 suggests undercorrection (increase K).\n\n")
  }
}
```

## 6.2 Genomic Offset (Climate Vulnerability)

```{r offset-section, results='asis'}
if (is.null(offset_df) || nrow(offset_df) == 0) {
  cat('<p class="unavailable">Genomic offset results not available. Run the Offset module with current + projected environmental data to estimate climate vulnerability.</p>\n')
} else {
  # offset_df from calc_genomic_offset() has columns:
  #   population, offset, offset_norm, lat, lon, risk_label, risk_colour
  off_tbl <- offset_df[order(offset_df$offset_norm, decreasing = TRUE), , drop = FALSE]

  display_cols <- intersect(c("population", "offset_norm", "offset", "risk_label"),
                             names(off_tbl))
  off_tbl <- off_tbl[, display_cols, drop = FALSE]

  col_renames <- c(
    population  = "Population",
    offset_norm = "Offset (normalised)",
    offset      = "Offset (raw)",
    risk_label  = "Climate risk"
  )
  names(off_tbl) <- col_renames[names(off_tbl)]

  print(knitr::kable(off_tbl, digits = 3,
    caption = paste0(
      "Genomic offset (Fitzpatrick & Keller 2015): RDA-based climate vulnerability. ",
      "Normalised offset 0 = lowest; 1 = highest within this dataset."
    )
  ))

  # Highlight high-risk populations
  n_severe <- sum(!is.na(offset_df$offset_norm) & offset_df$offset_norm >= 0.75)
  n_high   <- sum(!is.na(offset_df$offset_norm) &
                  offset_df$offset_norm >= 0.50 & offset_df$offset_norm < 0.75)

  if (n_severe > 0) {
    severe_pops <- offset_df$population[!is.na(offset_df$offset_norm) &
                                         offset_df$offset_norm >= 0.75]
    cat(sprintf(
      "\n**⚠️ %d population%s at severe climate exposure** (normalised offset ≥ 0.75): %s\n\n",
      n_severe, if (n_severe == 1) "" else "s",
      paste(severe_pops, collapse = ", ")
    ))
  }
  if (n_high > 0) {
    high_pops <- offset_df$population[!is.na(offset_df$offset_norm) &
                                       offset_df$offset_norm >= 0.50 &
                                       offset_df$offset_norm < 0.75]
    cat(sprintf(
      "**%d population%s at high climate exposure** (normalised offset 0.50–0.75): %s\n\n",
      n_high, if (n_high == 1) "" else "s",
      paste(high_pops, collapse = ", ")
    ))
  }

  cat('\n<p class="section-note">',
      'Genomic offset estimates the Euclidean distance in constrained ordination space ',
      'between a population\'s current and projected future environmental position. ',
      'Values are normalised to the maximum offset in this dataset. ',
      'Interpret alongside Ne and connectivity data — high offset in an already isolated ',
      'population represents compounded climate-change vulnerability.</p>\n')
}
```

---

# 7. Conservation Recommendations

```{r recommendations, results='asis'}
recs <- character(0)

# He-based
critical_pops <- df$population[!is.na(df$He) & df$He < 0.10]
high_risk_pops <- df$population[!is.na(df$He) & df$He >= 0.10 & df$He < 0.15]

if (length(critical_pops) > 0) {
  recs <- c(recs, sprintf(
    "**Urgent genetic rescue** for %s: He < 0.10 indicates critically low diversity and elevated extinction risk. Assess options for managed genetic exchange or ex-situ conservation.",
    paste(critical_pops, collapse = ", ")
  ))
}

if (length(high_risk_pops) > 0) {
  recs <- c(recs, sprintf(
    "**Genetic monitoring** for %s: He between 0.10–0.15. Establish regular monitoring; evaluate corridor restoration or translocation.",
    paste(high_risk_pops, collapse = ", ")
  ))
}

# Inbreeding
elevated_inb <- df$population[!is.na(df$inbreeding_f) & df$inbreeding_f > 0.1]
if (length(elevated_inb) > 0) {
  recs <- c(recs, sprintf(
    "**Inbreeding mitigation** for %s: F > 0.10. Consider pedigree management, or introduction of unrelated individuals from genetically distinct populations.",
    paste(elevated_inb, collapse = ", ")
  ))
}

# Ne
if (!is.null(ne_stats) && "ne" %in% names(ne_stats)) {
  ne_critical <- ne_stats$population[!is.na(ne_stats$ne) & ne_stats$ne < 50]
  if (length(ne_critical) > 0) {
    recs <- c(recs, sprintf(
      "**Priority Ne intervention** for %s: Ne < 50 means genetic drift will fix deleterious alleles rapidly. Immediate population augmentation or captive assurance colony should be evaluated.",
      paste(ne_critical, collapse = ", ")
    ))
  }
}

# FST connectivity
if (!is.null(fst_mat) && nrow(fst_mat) >= 2) {
  high_fst <- which(fst_mat > 0.25 & upper.tri(fst_mat), arr.ind = TRUE)
  if (nrow(high_fst) > 0) {
    recs <- c(recs,
      "**Connectivity assessment** recommended: One or more population pairs show very high F_ST (> 0.25), suggesting long-term isolation. Landscape modelling and corridor feasibility analysis are advised."
    )
  }
}

# Genomic offset / climate vulnerability
if (!is.null(offset_df) && nrow(offset_df) > 0 && "offset_norm" %in% names(offset_df)) {
  severe_pops <- offset_df$population[!is.na(offset_df$offset_norm) &
                                       offset_df$offset_norm >= 0.75]
  high_pops   <- offset_df$population[!is.na(offset_df$offset_norm) &
                                       offset_df$offset_norm >= 0.50 &
                                       offset_df$offset_norm <  0.75]

  if (length(severe_pops) > 0) {
    recs <- c(recs, sprintf(
      paste0(
        "**Climate-adaptive management** urgently needed for %s: normalised genomic offset ≥ 0.75 indicates a ",
        "large adaptive gap under projected climate change. Prioritise for assisted gene flow from ",
        "climatically pre-adapted source populations, ex-situ genetic banking, and evaluation for ",
        "assisted migration / managed relocation to suitable future climate space (where ecologically feasible)."
      ),
      paste(severe_pops, collapse = ", ")
    ))
  }
  if (length(high_pops) > 0) {
    recs <- c(recs, sprintf(
      paste0(
        "**Climate vulnerability monitoring** for %s: genomic offset 0.50–0.75. ",
        "Establish monitoring to track demographic responses to climate change; ",
        "model potential source populations for future assisted gene flow."
      ),
      paste(high_pops, collapse = ", ")
    ))
  }
}

if (length(recs) == 0) {
  recs <- c("No urgent genetic concerns identified. Maintain routine monitoring of diversity metrics. Preserve existing connectivity between populations.")
}

for (rec in recs) {
  cat(sprintf("- %s\n\n", rec))
}

cat('\n<p class="section-note">These recommendations are based on genetic data only. ',
    'Conservation decisions should integrate demographic surveys, habitat assessments, ',
    'and stakeholder consultation. Consult a qualified conservation geneticist before ',
    'implementing translocation or breeding programmes.</p>\n')
```

---

# 8. Methods & Citations

All analyses were performed with **PopGen Map v`r params$app_version`** ([github.com/csunbird/population-genomics-map](https://github.com/csunbird/population-genomics-map)).

**Diversity statistics:** Expected heterozygosity He = 2p(1−p) averaged across biallelic SNPs; nucleotide diversity π uses the Nei (1978) diploid-corrected estimator; inbreeding F = (He − Ho) / He. SNPs with > 50% missing data within a population are excluded.

**Population differentiation:** Pairwise F_ST estimated using the Hudson (1992) ratio-of-averages method. Threshold interpretation follows Wright (1978).

**Effective population size:** Estimated by the single-sample linkage disequilibrium method (Waples & Do 2008). IUCN/SSC thresholds from Frankham et al. (2014).

**Inbreeding:** F coefficient = (He − Ho) / He from VCF genotype data. For per-individual runs-of-homozygosity (F-ROH) analysis capturing recent inbreeding, use the F-ROH tab in the app. F-ROH threshold interpretation follows McQuillan et al. (2008).

**GEA:** Partial redundancy analysis (RDA) partials out geographic effects before testing environment-genotype associations. LFMM2 uses K latent factors to control for population structure (Caye et al. 2019). FDR correction by Benjamini-Hochberg.

**Genomic offset:** RDA-based climate vulnerability metric (Fitzpatrick & Keller 2015); Euclidean distance in constrained ordination space between current and projected environmental position.

### Key references

- Caye, K., et al. (2019). LFMM 2: fast and accurate inference of gene-environment associations in genome-wide studies. *Molecular Biology and Evolution*, 36(4), 852–860.
- Fitzpatrick, M.C. & Keller, S.R. (2015). Ecological genomics meets community-level modelling of biodiversity: mapping the genomic landscape of current and future environmental adaptation. *Ecology Letters*, 18, 1–16.
- Frankham, R., et al. (2014). Genetic Management of Fragmented Animal and Plant Populations. Oxford University Press.
- Hudson, R.R., Slatkin, M. & Maddison, W.P. (1992). Estimation of levels of gene flow from DNA sequence data. *Genetics*, 132, 583–589.
- Nei, M. (1978). Estimation of average heterozygosity and genetic distance from a small number of individuals. *Genetics*, 89, 583–590.
- Waples, R.S. & Do, C. (2008). LDNe: a program for estimating effective population size from data on linkage disequilibrium. *Molecular Ecology Resources*, 8(4), 753–756.
- Wright, S. (1978). Evolution and the Genetics of Populations, Vol. 4. University of Chicago Press.

---

<p style="font-size:0.8em; color:#aaa; text-align:center; margin-top:40px;">
PopGen Map v`r params$app_version` &nbsp;|&nbsp; MIT Licence &nbsp;|&nbsp; `r params$report_date`<br/>
For research and conservation planning use only. Not a substitute for expert conservation genetics advice.
</p>
